Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2017-09-30 LaTeX/Turnover.tex
working out general invasion criteria

We are particularly concerned with the conditions under which a rare neo-sex-determining allele increases in frequency, which occurs when the largest eigenvalue, $\lambda$, is greater than one. 
If the average change in frequency of the two haplotypes that carry the $m$ allele ($Am$ and $am$) is positive, invasion will always occur, i.e., if $(\lambda_{mA} + \lambda_{ma})/2 > 1$ then $\lambda > 1$. 
If neither haplotype increases in frequency ($\lambda_{mA}, \lambda_{ma} < 1$), the $m$ allele will not invade. 
Otherwise, the new sex-determining allele increases in frequency on one \textbf{A} background and declines on the other, and invasion requires 
\begin{equation}\label{eq:lambdasGen}
%\frac{\chi_{mA}}{\lambda_{mA}-1}+\frac{\chi_{ma}}{\lambda_{ma}-1}>0.
\chi_{ma}\left(\lambda_{mA}-1\right)+\chi_{mA}\left(\lambda_{ma}-1\right)>0.
\end{equation}
\noindent For example, if we assume that only the $mA$ haplotype has a positive growth rate ($\lambda_{ma}<1<\lambda_{mA}$), the second term on the left-hand side of \eqref{eq:lambdasGen} is negative and invasion requires that the growth rate of $mA$ haplotypes and the rate at which they are produced by recombination is sufficiently large relative to that of $ma$ haplotypes.
In other words, invasion requires that the average growth rate of the two haplotypes, weighted by the rates they are created by recombination, is positive.
%the number of recombinants from the fitter $mA$ haplotypes is larger than that of the less fit $ma$ haplotypes, relative to their growth rates.

>>>>

We are particularly concerned with the conditions under which a rare neo-sex-determining allele increases in frequency, which occurs when the largest eigenvalue, $\lambda$, is greater than one. 
%If the average change in frequency of the two haplotypes that carry the $m$ allele ($Am$ and $am$) is positive, invasion will always occur, i.e., if $(\lambda_{mA} + \lambda_{ma})/2 > 1$ then $\lambda > 1$. 
Given the characteristic polynomial and the Perron-Forbenius theorem, at least one $\lambda$ is greater than one when $b<2$ (see supplementary material for details), or  $(\lambda_{mA} - \chi_{mA})/2 + (\lambda_{ma} -  \chi_{ma})/2 > 1$.
In words, the neo-sex-determining allele will certainly spread when the average change in its frequency, considering that recombination breaks up mutant haplotypes but ignoring the fact that recombination also creates mutant haplotypes, is positive.
We are also guaranteed to have at least one $\lambda$ greater than one when 1 + b + c < 0.
This is never true if $\lambda_{mA} < 1$ and $\lambda_{ma} < 1$, i.e., if neither haplotype increases in frequency the neo-sex-determining allele will not invade. 
Otherwise, the new sex-determining allele increases in frequency on one \textbf{A} background and declines on the other, and invasion requires 
\begin{equation}\label{eq:lambdasGen}
%\frac{\chi_{mA}}{\lambda_{mA}-1}+\frac{\chi_{ma}}{\lambda_{ma}-1}>0.
\chi_{ma}\left(\lambda_{mA}-1\right)+\chi_{mA}\left(\lambda_{ma}-1\right)>0.
\end{equation}
\noindent For example, if we assume that only the $mA$ haplotype has a positive growth rate ($\lambda_{ma}<1<\lambda_{mA}$), the second term on the left-hand side of \eqref{eq:lambdasGen} is negative and invasion requires that the growth rate of $mA$ haplotypes and the rate at which they are produced by recombination is sufficiently large relative to that of $ma$ haplotypes.
In other words, invasion requires that the average growth rate of the two haplotypes, weighted by the rates they are created by recombination, is positive.
%the number of recombinants from the fitter $mA$ haplotypes is larger than that of the less fit $ma$ haplotypes, relative to their growth rates.
